\vspace{1.0cm} \huge \bf Supplemental information:\\ \LARGE Reassortment between influenza B lineages and the emergence of a co-adapted PB1-PB2-HA gene complex

### 2014-07-23 fluB.tex
Move speciesTree figure to supp.
Reduces number of figures in the main text.

Although the linkage between the two polymerase segments and HA is more difficult to explain, it is widely accepted that Victoria lineage HA had been restricted to eastern Asia between 1992 and 2000 \citep{nerome1998,shaw2002}, offering ample time for the budding Victoria lineage to accumulate alleles causing reassortment incompatibility.
Without more data it is difficult to tell how much of the linkage between PB1, PB2 and HA segments can be attributed to the population structure of influenza B viruses in the past.
However, there is some evidence that the linkage between PB1 and HA might not be a phenomenon restricted to influenza B viruses.
It has been established that at least for the 1957 and the 1968 influenza pandemics, caused by A/H2N2 and A/H3N2 subtypes, respectively, the viruses responsible were reassortants possessing PB1 and HA segments derived from avian influenza A viruses \citep{kawaoka1989}.
In addition, outdated techniques for producing vaccine seed strains through selection for HA-NA reassortants often yielded PB1-HA-NA reassortants as a side-effect \citep{bergeron2010,fulvini2011}.
Recent experiments have found that the presence or absence of a `foreign' PB1 segment can have dramatic effects on HA concentration on the surface of virions and total virion production \citep{cobbin2013}.
Bu we would also like to point out that there have been reassortant influenza A viruses circulating for prolonged periods of time in humans that did have disparate PB1 and HA segments, \textit{e.g.} H1N2 outbreaks in 2001 \citep{gregory2002} and H1N1/09 in 2009 \citep{smith2009}.

@@

\begin{figure}[h]
	\centering	
	\includegraphics[width=0.85\textwidth]{figures/InfB_methodSpeciesTree.pdf}
	\caption{\textbf{Calculating $\dtmrca$ from a species tree perspective.}
Consider an organism that has diverged into 3 taxa (A, B, C) with a genome comprised of 3 segments (blue, green and red).
Due to reassortments taxa A and B share a slightly more recent TMRCA in the red segment, likewise for taxa B and C in the green segment.
By comparing differences in TMRCAs between taxa A-B, A-C and B-C in blue, red and green segments we would find that the red segment has a lower `reassortment distance' (X) than the green segment (Y).
}
	\label{speciesTree}
\end{figure}

>>>>

Although the linkage between the two polymerase segments and HA is more difficult to explain, it is widely accepted that Victoria lineage HA had been restricted to eastern Asia between 1992 and 2000 \citep{nerome1998,shaw2002}, offering ample time for the budding Victoria lineage to accumulate alleles causing reassortment incompatibility.
Without more genomic data from the past, however, it is difficult to estimate to what extent influenza B virus population structure contributed to the development of the current segment linkage.
There is some evidence that the linkage between PB1 and HA might not be a phenomenon restricted to influenza B viruses.
It has been established that at least for the 1957 and the 1968 influenza pandemics, caused by A/H2N2 and A/H3N2 subtypes, respectively, the viruses responsible were reassortants possessing PB1 and HA segments derived from avian influenza A viruses \citep{kawaoka1989}.
In addition, outdated techniques for producing vaccine seed strains through selection for HA-NA reassortants often yielded PB1-HA-NA reassortants as a side-effect \citep{bergeron2010,fulvini2011}.
Recent experiments have found that the presence or absence of a `foreign' PB1 segment can have dramatic effects on HA concentration on the surface of virions and total virion production \citep{cobbin2013}.
Bu we would also like to point out that there have been reassortant influenza A viruses circulating for prolonged periods of time in humans that did have disparate PB1 and HA segments, \textit{e.g.} H1N2 outbreaks in 2001 \citep{gregory2002} and H1N1/09 in 2009 \citep{smith2009}.

@@

### 2014-07-23 fluB_supp.tex
Clean language per #38
Also add references to similar methods.
Move speciesTree figure to supp.
Reduces number of figures in the main text.

\begin{figure}
\centering  
\includegraphics[width=0.75\textwidth]  {supp_figures/InfB_Ntransitions.png}
\caption{\textbf{Numbers of trait transitions in each tree.}
The numbers shown are the mean inferred number of trait transitions (minus one to account for the initial Vic--Yam split) in a given tree and trait combination.
The numbers in brackets correspond to 95\% highest posterior density intervals.}
\label{Ntransitions}
\end{figure}

@@

>>>>

\begin{figure}
\centering  
\includegraphics[width=0.75\textwidth]  {supp_figures/InfB_Ntransitions.png}
\caption{\textbf{Numbers of trait transitions in each tree.}
The numbers shown are the mean inferred number of trait transitions (minus one to account for the initial Vic--Yam split) in a given tree and trait combination.
The numbers in brackets correspond to 95\% highest posterior density intervals.
Transitions may not be independent when more than one segment reassorts at the same time.}
\label{Ntransitions}
\end{figure}

@@

\begin{figure}[h]
	\centering	
	\includegraphics[width=0.85\textwidth]{figures/InfB_methodSpeciesTree.pdf}
	\caption{\textbf{Calculating $\dtmrca$ from a species tree perspective.}
Consider an organism that has diverged into 3 taxa (A, B, C) with a genome comprised of 3 segments (blue, green and red).
Due to reassortments taxa A and B share a slightly more recent TMRCA in the red segment, likewise for taxa B and C in the green segment.
By comparing differences in TMRCAs between taxa A-B, A-C and B-C in blue, red and green segments we would find that the red segment has a lower `reassortment distance' (X) than the green segment (Y).
}
	\label{speciesTree}
\end{figure}
